---
title: "Zoom in on Granulocytes"
date: University of Bern, DBMR, Visceral and Transplantation Surgery lab
author: "Tess Brodie"
output: 
  html_document:
    toc_depth: 2
---


```{r setup, include = FALSE, dev='png'}
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
knitr::opts_chunk$set(autodep = TRUE, cache = TRUE, cache.lazy = FALSE)
knitr::opts_chunk$set(dev = 'png')
```

## Load packages

```{r load-libs, message = FALSE, warning = FALSE}
library(CATALYST)
library(dplyr)
library(workflowr)
library(flowCore)
library(ggplot2)
library(diffcyt)
library(cowplot)
library(viridis)
library(RColorBrewer)
```

### Load data

```{r load-data}
sce <- readRDS(file.path("output", "sce.rds"))
```

### Filter on Granulocytes 

```{r Filter_BMDM}
sceG <- filterSCE(sce, k = "mm",
               cluster_id %in% "Granulocytes")

plot_grid(
  plotDR(sce, dr= "TSNE", color_by = "mm"),
  plotDR(sceG, dr= "TSNE", color_by = "mm"))
```

### Target expression in Granulocytes over time

```{r ExprHeatmap_granulos}
plotExprHeatmap(sceG,
    row_anno = FALSE,   # don't annotate samples
    row_clust = FALSE,  # keep samples in original order
    col_clust = TRUE,  # keep markers in original order
    bin_anno = FALSE,   # don't annotate bins
    bars = FALSE,       # don't include sample sizes
    scale = "last")     # aggregate, then scale
```

## Number of Granulocytes per sample

```{r Granulo_count}
plotCounts(sceG, 
           group_by = "sample_id", 
           color_by = "condition")
```

## TSNE of Granulocytes colored by condition

```{r TSNE_condition}
sceG <- runDR(sceG, dr = "TSNE", features = NULL)
plotDR(sceG, dr = "TSNE", color_by = "condition")
```

## Facet by condition

```{r TSNE_condition_facet}
plotDR(sceG, dr = "TSNE", color_by = "condition", facet_by = "condition")
```

## Granulocyte Type markers TSNE facet

```{r plotDR-expression_Type, fig.width = 12, fig.height = 5}
p <- plotDR(sceG, dr = "TSNE", color_by = type_markers(sce))
p$facet$params$ncol <- 4; p
```

## ...split by condition {.tabset}

```{r plotDR-expression-condition, results = "asis", fig.width = 12, fig.height = 5}
for (i in state_markers(sce)) {
    p <- plotDR(sceG, dr = "TSNE", color_by = i, facet_by = "condition")
    cat("### ", i, "\n"); print(p); cat("\n\n")
}
```

## Cluster Granulocytes into 3 subsets
```{r cluster_3, message = FALSE}
sceG <- cluster(sceG, maxK = 3)
```

## Expression heatmap of 3 Granulocyte subsets
```{r plotExprHeatmap_5, fig.width = 6, fig.height = 5}
plotExprHeatmap(sceG, features = type_markers(sce), 
    by = "cluster_id", k = "meta3", row_anno = FALSE, bars = TRUE)
```

## 3 Granulocyte clusters overlaid on TSNE plot
```{r TSNE_meta5}
plotDR(sceG, dr = "TSNE", color_by = "meta3")
```

## Granulocyte frequency by cluster

```{r Granulos_Freqheatmap_cluster}
plotFreqHeatmap(sceG, 
    k = "meta3",
    hm_pal = rev(hcl.colors(10, "RdBu")),
    k_pal = hcl.colors(7, "Zissou 1"),
    col_clust = FALSE, col_anno = FALSE,
    row_anno = FALSE,
    bars = TRUE, perc = TRUE)
```

## Differential state Analysis for 3 Granulocyte clusters
```{r DS}
# create design matrix
design <- createDesignMatrix(ei(sceG), "condition")
colnames(design) <- gsub("condition", "", colnames(design))

# run DS analysis for control vs. each timepoint
for (i in colnames(design)[-1]) {
  # create contrast matrix
  contrast <- createContrast(as.numeric(colnames(design) == i))
    
  # run DS analysis & extract results table
  res_DS <- diffcyt(sceG, clustering_to_use = "meta3",
                  analysis_type = "DS", method_DS = "diffcyt-DS-limma",
                  design = design, contrast = contrast, verbose = FALSE)

  # extract results table
  tbl_DS <- rowData(res_DS$res)

  # write table to .csv
    fn <- file.path("output", sprintf("DS-resultsGranulos_%s.csv", i))
    write.csv(tbl_DS, fn, row.names = FALSE)
    
  # subset control and timepoint of interest
    sub <- filterSCE(sceG, condition %in% c("control", !!i))
    
  # plot differential heatmap
    p <- plotDiffHeatmap(sub, tbl_DS, top_n = 30, all = TRUE, col_anno = "condition")
    cat("## ", i, "\n"); print(p); cat("\n\n")
} 
```

## Differential Abundances analysis with 3 Granulocytes subsets

# DA analysis {.tabset}
```{r da, results = "asis", fig.width = 8, fig.height = 6}
# create design matrix
design <- createDesignMatrix(ei(sceG), "condition")
colnames(design) <- gsub("condition", "", colnames(design))

# run DA analysis for control vs. each timepoint
for (i in colnames(design)[-1]) {
    # create contrast matrix
    contrast <- createContrast(as.numeric(colnames(design) == i))
    
    # run DA analysis
    res <- diffcyt(sceG, clustering_to_use = "meta3",
        analysis_type = "DA", method_DA = "diffcyt-DA-edgeR",
        design = design, contrast = contrast, verbose = FALSE)
    
    # extract results table
    tbl <- rowData(res$res)
    
    # write table to .csv
    fn <- file.path("output", sprintf("DA-resultsGranulos_%s.csv", i))
    write.csv(tbl, fn, row.names = FALSE)
    
    # subset control and timepoint of interest
    sub <- filterSCE(sceG, condition %in% c("control", !!i))
    
    # plot differential heatmap
    p <- plotDiffHeatmap(sub, tbl, all = TRUE, col_anno = "condition")
    cat("## ", i, "\n"); print(p); cat("\n\n")
}
```



